Capturing the shape of our experiences and how we remember them

### 2018-08-20 paper/main.tex
edits to results

\subsubsection{Fitting the topic model to the video text and recall transcripts}
A topic model was used to estimate the most likely mixture of topics for a given sample of text. First, the video was manually segmented into 1000 scenes, and a collection of descriptive features was manually transcribed. For each scene, we considered the following features: scene details (i.e. a sentence or two describing what happened in that scene, space (indoor or outdoor), name of all the characters in the scene, name of the character in focus, name of the character speaking, location, camera angle, music presence, and words on the screen. We concatenated the text for all of these features within each segment, creating a 'bag of words' describing each scene. We then transformed the text descriptions into overlapping windows of 50 scene segments. For example, the first text sample comprised the text from the first 50 segments, the second comprised the text from n+1:n+51, and so on. We trained our model using these overlapping text samples using scikit-learn's (version 0.19.1) `CountVectorizer' and `LatentDirichletAllocation' classes.  First, the text was transformed into a vector of word counts (default parameters). This gave a word count vector for each scene in the video.  Then, the word count vectors were used to fit a topic model (topics=100, method=batch). We transformed the text descriptions using the model resulting in a scenes (1000) by topics (100) matrix. The scene descriptions often spanned multiple timepoints (i.e. TRs). To account for this, we expanded the video model by copying the rows of the model for as many timepoints that the scene description spanned. After this expansion, the shape of the model was the length of the duration of the video (1976 TRs).

@@

\subsubsection{Choosing topic model parameters}
There were 3 critical parameters related to fitting the topic model: 1) the number of topics, 2) the window size of text descriptions of the video used to fit the model, and 3) the window size of recall sentences used to transform the recall data.  To chose these parameter values, we performed a grid search where the range of possible parameter values was 1, 5, 10, 25, 50, 100, 200, and 500. Our optimization objective was defined as the correlation between the hand annotated memory performance and the root mean squared distance between the video model and the recall model before any further processing (e.g. hidden Markov modelling, averaging within event, etc). While many of the parameter combination elicited moderately high correlations, the optimal choice was 100 topics, 50 video segments and 10 recall sentences.

\subsubsection{Extracting events using a hidden Markov model}
The topic model timepoint-by-timepoint correlation matrices all exhibited a block-diagonal structure (with small off-diagonal values), suggesting that the models were comprised of a number of sequential `states' (or events). To capture this structure, we fit the video and each recall model using a hidden Markov model (HMM). Given a number of states or events ($k$), the HMM recovers a set of labels that segments consecutive timepoints into $k$ events ~\citep{Rabi89, BaldEtal17}. To implement this analysis, we used the Brainiak toolbox \citep{BaldEtal17}.

Our metric for choosing the `best fitting' HMM was to choose the model with the $k$ value that maximized the ratio of the average `within-event' correlation values (i.e. the correlation values for blocks of consecutive timepoints the model identified as one event) and the average `across-event' correlation (i.e. the rest of the correlation values). Additionally, we included a penalty parameter that was proportional to the smoothing of the model that preferred models with smaller $k$ values. We chose $k$ values separately for the video model and for each recall model.  Then, using the best $k$ values, We fit a separate HMM for each topic model. Finally, we averaged over timepoints identified to be in the same event resulting in a events by topics matrix for the video model and each of the recall models.
%
\subsubsection{Matching recall events to video events}
To figure out which video event each recall event referred to, we correlated the video events model and each recall events model. This resulted in a video events (34) by recall events (8-27) correlation matrix (for each participant) which contains the similarity between each video event and each recall event.  To find the most likely video event that a given recall event referred to, we computed the argmax over the columns of this matrix.  The result was a list of video event indices for each participant. These indices are analogous to the values found in a ``recall matrix'' from a free recall list learning experiment, but represent the recall of particular events (instead of words, for example).

\subsubsection{Visualizing the video and recall event models}
To better understand the temporal structure of the video event model (34 events by 100 topics) and the recall event models (8-27 events by 100 topics), we used a technique called UMAP ~\citep{McInHeal18} to reduce the ``topic-space'' from 100 dimensions down to 2 dimensions. UMAP is a nonlinear dimensionality reduction technique which models the manifold of the data with a fuzzy topological structure, and then searches for a (2D) projection of the data that has the closest equivalent fuzzy topological structure. We concatenated (vertically stacked) all event models (video, average recall, and individual recall), and then fit and transformed all of the models at once. This assured that the models were projected into the same space.

\subsubsection{Vector field analysis}
To quantify (and visualize) the flow of recall from event to event, we performed a vector field analysis.  We tiled the 2D topic space (x, y: -6 to 6 by .25) with an evenly spaced grid. For each grid point, we drew a square around the point (height/width=.5). Then, we tested whether any line segments (formed by event recall transitions) passed through this area of the topic space.  For example, say that a participant transitioned from recalling event 2 to event 3. These 2 recall events correspond to 2 points in topic space, and connecting them forms a line segment. We collected all line segments that passed through a given section of topic space (collapsing across participants). To plot the average direction of the line segments (i.e. the arrows for each grid point in ~\ref{fig:trajectory}b), we converted each of them to unit vectors and then averaged. For grid points where the direction was consistent (across all participants contributing to that point), the length of the arrow approaches 1, whereas if the direction was random the length of the arrow approaches 0. Lastly, we converted each unit vector to an angle (in radians) by taking the inverse tangent of the x, y components of the vector. To test whether the distribution of angles was significantly non-uniform (i.e. displayed a preferred direction across participants), we performed a Rayleigh test on the angles. Arrows where the Rayleigh test was significant are displayed in color while non-significant tests are displayed in gray with lower opacity.

>>>>

\subsubsection{Fitting the topic model to the video text and recall transcripts}
A topic model was used to estimate the most likely mixture of topics for a given sample of text. First, the video was manually segmented into 1000 scenes, and a collection of descriptive features was manually transcribed. For each scene, we considered the following features: scene details (i.e. a sentence or two describing what happened in that scene), space (indoor or outdoor), name of all the characters in the scene, name of the character in focus, name of the character speaking, location, camera angle, music presence, and words on the screen. We concatenated the text for all of these features within each segment, creating a ``bag of words'' describing each scene. We then transformed the text descriptions into overlapping windows of 50 scene segments. For example, the first text sample comprised the text from the first 50 segments, the second comprised the text from n+1:n+51, and so on. We trained our model using these overlapping text samples using scikit-learn's (version 0.19.1) `CountVectorizer' and `LatentDirichletAllocation' classes.  First, the text was transformed into a vector of word counts (after removing English stopwords). This gave a word count vector for each scene in the video.  Then, the word count vectors were used to fit a topic model (topics=100, method=`batch'). We transformed the text descriptions using the model resulting in a scenes (1000) by topics (100) matrix. The scene descriptions often spanned multiple timepoints (i.e. TRs). To account for this, we expanded the video model by copying the rows of the model for as many timepoints that the scene description spanned. After this expansion, the shape of the model was the length of the duration of the video (1976 TRs).

@@

\subsubsection{Choosing topic model parameters}
There were 3 critical parameters related to fitting the topic model: 1) the number of topics, 2) the window size of text descriptions of the video used to fit the model, and 3) the window size of recall sentences used to transform the recall data.  To chose these parameter values, we performed a grid search where the range of possible parameter values was 1, 5, 10, 25, 50, 100, 200, and 500. Our optimization objective was defined as the correlation between the hand annotated memory performance and the root mean squared distance between the video model and the recall model before any further processing (e.g. hidden Markov modelling, averaging within event, etc). While many of the parameter combination elicited moderately high correlations, the optimal choice was 100 topics, 50 video segments and 10 recall sentences.
%NOTE: I did run this analysis, but before that i arbitrarily selected some parameters that ended up being very close to the best.  The reason I bring this up is that the correlations between some of the memory metrics (k, precision, distinctiveness) and hand annotated accuracy could be biased because we searched for the optimal parameters. Its not `the same' measure, but they are related to each other, so we'll have to be extra careful not to over interpret the correlations in this study.

\subsubsection{Extracting events using a hidden Markov model}
The topic model timepoint-by-timepoint correlation matrices all exhibited a block-diagonal structure (with small off-diagonal values), suggesting that the models were comprised of a number of sequential `states' (or events, e.g. Supp. Fig.~\ref{fig:corrmats}). To capture this structure, we fit the video and each recall model using a hidden Markov model (HMM). Given a number of states or events ($k$), the HMM recovers a set of labels that segments consecutive timepoints into $k$ events ~\citep{Rabi89, BaldEtal17}. To implement this analysis, we used the Brainiak toolbox \citep{BaldEtal17}.

Our metric for choosing the ``best fitting'' HMM was to choose the model with the $k$ value that maximized the ratio of the average `within-event' correlation values (i.e. the correlation values for blocks of consecutive timepoints the model identified as one event) and the average `across-event' correlation (i.e. the rest of the correlation values). Additionally, we included a penalty parameter that was proportional to the smoothing of the model that preferred models with smaller $k$ values. We chose $k$ values separately for the video model and for each recall model.  Then, using the best $k$ values, We fit a separate HMM for the video and each recall model. Finally, we averaged over timepoints identified to be in the same event resulting in a events by topics matrix for the video model and each of the recall models.

\subsubsection{Matching recall events to video events}
To estimate which video event each recall event referred to, we correlated the video events model and each recall events model. This resulted in a video events (34) by recall events (8-27) correlation matrix (for each participant) which contains the similarity between each video event and each recall event.  To find the most likely video event that a given recall event referred to, we computed the argmax over the columns of this matrix.  The result was a list of video event indices for each participant. These indices are analogous to the values found in a ``recall matrix'' from a free recall list learning experiment, but represent the recall of particular events (instead of words, for example).

\subsubsection{Visualizing the video and recall event models}
To visualize the temporal structure of the video event model (34 events by 100 topics) and the recall event models (8-27 events by 100 topics), we used a technique called UMAP ~\citep{McInHeal18} to reduce the ``topic-space'' from 100 dimensions down to 2 dimensions. UMAP is a nonlinear dimensionality reduction technique which models the manifold of the data with a fuzzy topological structure, and then searches for a (2D) projection of the data that has the closest equivalent fuzzy topological structure. We concatenated (vertically stacked) all event models (video, average recall, and individual recall), and then fit and transformed all of the models at once. This assured that the models were projected into the same space.

\subsubsection{Vector field analysis}
To quantify the flow of recall from event to event, we performed a vector field analysis.  We tiled the 2D topic space (x, y: -6 to 6 by .25) with an evenly spaced grid. For each grid point, we drew a circle around the point (radius=.5). Then, we tested whether any line segments (formed by event recall transitions) passed through this area of the topic space.  For example, say that a participant transitioned from recalling event 2 to event 3. These 2 recall events correspond to 2 points in topic space, and connecting them forms a line segment. We collected all line segments that passed through a given section of topic space (collapsing across participants). To plot the average direction of the line segments (i.e. the arrows for each grid point in ~\ref{fig:trajectory}b), we converted each of them to unit vectors and then averaged. For grid points where the direction was consistent (across all participants contributing to that point), the length of the arrow approaches 1, whereas if the direction was random the length of the arrow approaches 0. Lastly, we converted each unit vector to an angle (in radians) by taking the inverse tangent of the x, y components of the vector. To test whether the distribution of angles was significantly non-uniform (i.e. displayed a preferred direction across participants), we performed a Rayleigh test on the angles (p$<$.001, FDR-corrected at p$<$.05 using Benjamani-Hochberg procedure). Arrows where the Rayleigh test was significant are displayed in color while non-significant tests are displayed in gray with lower opacity.
